Nitric-oxide-feedback to ciliary photoreceptor cells orchestrates UV avoidance

### 2022-10-03 text/Manuscript.Rmd
Add correlation data

To test if NO is produced during UV/violet responses, we injected zygotes with the fluorescent NO-reporter DAF-FM and imaged two-day-old larvae during the local stimulation of the cPRC cilia by 405 nm light (Figure 4---figure supplement 1). Following light stimulation, we detected an increase in DAF-FM fluorescence in the region of the INNOS cells. 
We next analysed the dynamics of the cPRC circuit following light stimulation. As we have shown previously [@Veraszto2018], a 20-sec illumination of the cPRC ciliary region lead to a decrease in cPRC calcium levels, likely due to the hyperpolarisation of the cPRCs. After 20 sec, calcium levels in cPRCs were raising again, reaching higher levels than at the start of the stimulus - a response that may involve depolarisation. This activation phase occurs after the end of the stimulation and may be due to a slower neuroendocrine feedback [@Veraszto2018]. 
To determine whether NO mediates such a feedback, we measured calcium signals in *NOS*-mutant larvae. While we detected the inhibition phase in cPRC in *NOS*-mutants, this was not followed by activation and calcium levels remained low. These results revealed an essential requirement for NOS signalling in the second phase of the cPRC response.

We next quantified the responses of the INNOS and INRGW interneurons following a UV/violet stimulus to the cPRCs. To unambiguously identify the cells from which we recorded calcium signals, we developed an on-slide immunostaining method. We used the cell-specific neuropeptide markers anti-RYamide (INNOS) and anti-RGWa (INRGW) [@Conzelmann2011] to mark the interneurons. Based on the position of the nuclei, we could correlate live and fixed samples at a single-cell precision (Figure 4---figure supplement 2). In both wild type and *NOS*-mutant larvae, INNOS cells showed increase calcium signals during the inhibition phase of the cPRCs. In contrast, calcium signals increased in INRGW cells in wild-type larvae during the second activation phase of cPRCs, a response that was lacking in *NOS*-mutants. This suggests that there is no circuit output to the Ser-h1 cells and the ciliary band in *NOS*-mutants, consistent with the observed behavioural defects.

>>>>

To test if NO is produced during UV/violet responses, we injected zygotes with the fluorescent NO-reporter DAF-FM and imaged two-day-old larvae during the local stimulation of the cPRC cilia by 405 nm light (Figure 4---figure supplement 1). Following light stimulation, we detected an increase in DAF-FM fluorescence in the region of the INNOS cells (Figure 4A). 
We next analysed the dynamics of the cPRC circuit following light stimulation. As we have shown previously [@Veraszto2018], a 20-sec illumination of the cPRC ciliary region lead to a decrease in cPRC calcium levels, likely due to the hyperpolarisation of the cPRCs. After 20-sec, calcium levels in cPRCs were raising again, reaching higher levels than at the start of the stimulus - a response that may involve depolarisation. This activation phase occurs after the end of the stimulation and may be due to a slower neuroendocrine feedback [@Veraszto2018]. To determine whether NO mediates such a feedback, we measured calcium signals in *NOS*-mutant larvae. While we detected the inhibition phase in cPRC in *NOS*-mutants, this was not followed by activation and calcium levels remained low (Figure 4B and C). These results revealed an essential requirement for NOS signalling in the second phase of the cPRC response.
We next quantified the responses of the INNOS and INRGW interneurons following a UV/violet stimulus to the cPRCs. To unambiguously identify the cells from which we recorded calcium signals, we developed an on-slide immunostaining method. We used the cell-specific neuropeptide markers anti-RYamide (INNOS) and anti-RGWa (INRGW) [@Conzelmann2011] to mark the interneurons. Based on the position of the nuclei, we could correlate live and fixed samples at a single-cell precision (Figure 4---figure supplement 2). In both wild type and *NOS*-mutant larvae, INNOS cells showed increase calcium signals during the inhibition phase of the cPRCs (Figure 4D and E). In addition, In-NOS depolarisation was suppressed in NIT-GC2 molphants where cPRC hyperpolarisation was suppressed. This indicates that cPRC hyperpolarisation is required for IN-NOS depolarisation (Figure 4F). In contrast, calcium signals increased in INRGW cells in wild-type larvae during the second activation phase of cPRCs, a response that was lacking in *NOS*-mutants (Figure 4G and H). This suggests that there is no circuit output to the Ser-h1 cells and the ciliary band in *NOS*-mutants, consistent with the observed behavioural defects(Figure 4I).
